---
title: "Exploring patterins in pottery use in the southwestern Taiwan during the Neolithic and Iron Age transition "
author:
  - Li-Ying Wang:
      correspondence: "yes"
      email: liyingw15@gmail.com
      orcid: 0000-0001-9060-2125
      institute: Academia Sinica
  - XX XX:
      institute: National Taiwan University
  - YY YY :
      institute: Academia Sinica
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

# Introduction

The transition from Neolithic period to Iron Age was often accompanied with some changes in activities related to food production or cultural practices due to the introduction of new tools. In southwestern Taiwan, archaeological materials from the late Neolithic period, referred as the Dahu culture, show distinct cultural elements compared to those from the Iron Age, Niaosung culture. Although local variations present across southwestern Taiwan, the differences between those two prehistoric cultures are generally evident in ceramic styles and burial practices. For examples, the Dahu culture is characterized by the use of gray-black pottery produced through reduction firing, a technique absent in the Niaosung assemblages, which instead predominated by red pottery normally through oxidation production. As for burial practices, prone and urn burials were common in the Dahu culture, while extended burials were the major practice in Niaosung culture. It remains unclear whether these cultural changes resulted from the coexistence or succession of two distinct cultural groups, or from cultural changes within the same group because of the adoption of new technologies or cultural elements.

Studying the dietary practices and survival strategies adopted by these prehistoric groups in the same environment will help us understand whether these different cultural groups had unique natural resource utilization practices or were deeply influenced by environmental adaptations. This project aims to address these research questions by focusing on materials from archaeological sites in Tainan, including Sanbaopi, Wushantou, and Shiqiao. Those sites cover a chronological range from the late Neolithic Dahu period to the Iron Age, making it an ideal location for studying dietary practices and cultural transformations. For example, Sanbaopi site, due to multiple excavations during the construction of the Tainan Science Park, has yielded abundant prehistoric artifacts and features, including ash pits, hearths, and burials, indicating long-term habitation.

To investigate prehistoric diets, this research will focus on the analysis of organic residues in pottery, particularly the dietary content reflected in lipids. These findings will be compared with evidence from faunal remains and other artifacts. In addition to lipid analysis of pottery samples, lipid analysis will also be conducted on common animal bone specimens from the sites, such as deer and pig remains. This will help establish a more detailed database of animal fatty acid isotope values specific to the region for comparison with the values obtained from pottery samples. The goal is to gain a deeper understanding of prehistoric cultural changes and environmental adaptations in Taiwan.


# Background

# Methods

# Results

<!-- Here's some example analysis code: -->

```{r}
#| label: get-data
#| eval: false
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r}
#| label: fig-demo-plot
#| fig-cap: "A plot of random numbers"
plot(rnorm(10))
```

@fig-demo-plot shows how we can have a caption and cross-reference for a plot. Note that figure label and cross-references must both be prefixed with `fig-`

```{r}
#| label: demo-inline-code
#| echo: false
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```
